Inferring Continuous and Discrete Population Genetic Structure Across Space

### 2017-08-18 writeup/conStruct.tex
intro changes

A pattern of isolation by distance (IBD) arises when migration in a population is spatially limited \citep{Wright1943},
and has been modeled in both spatially continuous frameworks \citep{malecot1969mathematics,Slatkin1985}, 
in which an individual's parents are found within some radius of that individual's location,
and discrete, ``stepping-stone'' frameworks \citep{kimura1964stepping},
in which migration occurs only between neighboring demes.
Given the generality of the circumstances that generate a pattern of isolation by distance, 
it is unsurprising that IBD is found everywhere in nature \citep{meirmans2012,Sexton_etal_2014}.

The ubiquity of isolation by distance presents a challenge for models of discrete population structure,
as it is frequently difficult to determine whether observed patterns of genetic variation are 
continuously distributed across a landscape, or instead are partitioned in discrete clusters.
This problem can be compounded if sampling is done unevenly or discretely across a population or species' range,
and has given rise to a debate in the population genetic literature about whether 
individuals are best described as being drawn from continuous clines or discrete clusters 
\citep[e.g.][]{SerrePaabo2004,rosenberg2005clines}.
As a further complication, if local genetic drift is strong,
continous models are expected to give rise to spatially patchy patterns
\citep{SLFV,bramson1989crabgrass}.

Existing model-based clustering approaches can only describe continuous patterns of variation using
discrete clusters, 
and so tend to erroneously describe continuously distributed variation with multiple clusters that 
show spatially autocorrelated cluster membership \citep{Frantz2009,meirmans2012}.
In analyses of empirical datasets, which often are characterized by IBD,
model-based clustering approaches will therefore tend to overestimate the number of discrete clusters present,
with downstream implications for interpretation of results.
For example, conservation resources might be spread too thinly over areas 
that are spuriously assumed to be discrete management units.
Or, in studies of locally adaptive loci, or of the genetics of disease,
discrete populations may be modeled in separately,
reducing the global power to detect the genomic basis of complex phenotypes.

>>>>

A pattern of isolation by distance (IBD) arises when migration in a population is spatially limited \citep{Wright1943},
and has been modeled in both spatially continuous frameworks \citep{malecot1969mathematics,Slatkin1985}, 
in which an individual's parents are found within some radius of that individual's location,
and discrete, ``stepping-stone'' frameworks \citep{kimura1964stepping},
in which migration occurs only between neighboring demes.
Given the generality of the circumstances that generate a pattern of isolation by distance, 
it is unsurprising that IBD is very widespread in nature \citep{meirmans2012,Sexton_etal_2014}.

The ubiquity of isolation by distance presents a challenge for models of discrete population structure,
as it is frequently difficult to determine whether observed patterns of genetic variation are 
continuously distributed across a landscape, or instead are partitioned in discrete clusters.
This problem can be compounded if sampling is done unevenly or discretely across a population or species' range,
and has given rise to a debate in the population genetic literature about whether 
individuals are best described as being drawn from continuous clines or discrete clusters 
\citep[e.g.][]{SerrePaabo2004,rosenberg2005clines}.
%As a further complication, if local genetic drift is strong,
%continous models are expected to give rise to spatially patchy patterns
%\citep{SLFV,bramson1989crabgrass}.

Existing model-based clustering approaches can only describe continuous patterns of variation using
discrete clusters, and so tend to erroneously describe continuously distributed variation with multiple clusters that 
show spatially autocorrelated cluster membership \citep{Frantz2009,meirmans2012}.
In analyses of empirical datasets, which often are characterized by IBD,
model-based clustering approaches will therefore tend to overestimate
the number of discrete clusters present.

%with downstream implications for interpretation of results.
%For example, conservation resources might be spread too thinly over areas 
%that are spuriously assumed to be discrete management units.
%Or, in studies of locally adaptive loci, or of the genetics of disease,
%discrete populations may be modeled in separately,
%reducing the global power to detect the genomic basis of complex phenotypes.
